r5r: rapid realistic routing on multimodal transport networks with R5 in R

### 2021-01-18 r-package/paper2020/paper2020.Rmd
add referecen to Levinson et al, plus text edits

Transport routing is the process of finding alternative routes and costs that connect places on a given transport network, and it is a key step in transport accessibility analysis, fleet allocation and transport simulation and planning more broadly. Researchers and practitioners still frequently face practical challenges when carrying out routing analysis tasks, such as the monetary costs of licensed software, limited data availability for data hungry models, and the long computation times required to run multiple routing scenarios, particularly in large and complex multimodal transport networks.

While there is a growing number of open source routing models [@opentripplanneropentripplanner; @lovelace2019stplanr; @padgham2019dodgr], most options available do not efficiently process large transport networks. This paper presents [r5r](https://ipeagit.github.io/r5r/), a new open source R package for routing on multimodal transport networks based on the [Rapid Realistic Routing on Real-world and Reimagined networks (R<sup>5</sup>)](https://github.com/conveyal/r5). R<sup>5</sup> is a powerful next-generation routing engine written in Java and developed at Conveyal [@conway2017evidencebased; @conway2018accounting] to provide an efficient backend for analytic applications - such as accessibility assessment. The r5r package provides a simple and friendly interface to run R<sup>5</sup> locally from within R, which allows users to efficiently calculate travel time matrices or generate multiple route alternatives between origins and destinations using seamless parallel computing.

<!-- More importantly, those tools generate deterministic estimates of travel times without accounting for variability in results (due to service reliability issues, for example) and only return the fastest route option between a set of origins and destinations (see [@conway2018accounting] for a more in-depth discussion). This simple approach overlooks how often people's actual route choices deviate from shortest travel time paths, what can occur due to multiple factors such as imperfect information, monetary costs and safety issues [@larsen2011travel; @zhu2015people; @conway2019getting]. -->

@@

The r5r package has low data requirements and is easily scalable, allowing fast computation of routes and travel times at either the city or country levels of analysis. Street networks from [OpenStreetMap](https://www.openstreetmap.org/) (OSM) are automatically combined with one or multiple public transport data sets provided by the user in the standard [General Transit Feed Specification](https://developers.google.com/transit/gtfs/) (GTFS) format to create a routable transport network.

@@

<!-- Rafa's notes -->

<!-- jean: #>-  mass of assumptions regarding decision-making and routing (Miller, 2018).  -->

<!-- Haugen, 2011; Schwanen, 2008 -->

<!-- Miller, E.J., 2018. Accessibility: measurement and application in transportation planning. Transp. Rev. 38, #>551–555. https://doi.org/10.1080/01441647.2018.1492778 -->

<!-- This article presents **r5r**, a new open-source `R` package for rapid  -->

<!-- realistic routing on multimodal transport networks. -->

>>>>

Transport routing is the process of finding the alternative routes and costs that connect places on a given transport network, and it is a key step in transport accessibility analysis, fleet allocation and transport simulation and planning more broadly [@levinson2020manual]. Researchers and practitioners still frequently face practical challenges when carrying out routing analysis tasks due to the monetary costs of licensed software, limited data availability for data hungry models, and the long computation times required to run multiple routing scenarios, particularly in large and complex multimodal transport networks.

While there is a growing number of open source routing models [@opentripplanneropentripplanner; @lovelace2019stplanr; @padgham2019dodgr], most options available do not efficiently process large transport networks. This paper presents [r5r](https://ipeagit.github.io/r5r/), a new open source R package for routing on multimodal transport networks based on the [Rapid Realistic Routing on Real-world and Reimagined networks (R<sup>5</sup>)](https://github.com/conveyal/r5). R<sup>5</sup> is a powerful next-generation routing engine written in Java and developed at Conveyal [@conway2017evidencebased; @conway2018accounting] to provide an efficient backend for analytic applications - such as accessibility analysis. The r5r package provides a simple and friendly interface to run R<sup>5</sup> locally from within R, which allows users to efficiently calculate travel time matrices or generate multiple route alternatives between origins and destinations using seamless parallel computing.

<!-- More importantly, those tools generate deterministic estimates of travel times without accounting for variability in results (due to service reliability issues, for example) ... 666666666666

@@

The r5r package has low data requirements and is easily scalable, allowing fast computation of routes and travel times at either city or country-level analysis. It creates a routable transport network using street network data from [OpenStreetMap](https://www.openstreetmap.org/) (OSM), and it  automatically combines public transport data if they are provided by the user in the standard [General Transit Feed Specification](https://developers.google.com/transit/gtfs/) (GTFS) format .

@@
